Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome

### 2023-05-16 paper/paper.qmd
Reorganizd paper subdir
Enforce refs before figures
Add performance figure & placeholders

---
title: "Severe CDI manuscript"
bibliography: references.bib
fontsize: 11pt
date: last-modified
date-format: medium
format: 
  manuscript-pdf:
    keep-tex: true
    include-in-header: header.tex
    linenumbers: true
    doublespacing: true
    lof: false
  #gfm: default
  html: default
runningtitle: Predicting Severe _C. difficile_ infections
runningauthor: Sovacool
papertype: Preprint
author:
  - name: Kelly L. Sovacool
    orcid: 0000-0003-3283-829X
    affiliations:
      - ref: bioinf
  - name: Sarah E. Tomkovich
    affiliations:
      - ref: micro
  - name: Megan L. Coden
    affiliations:
      - ref: mcdb
  - name: Vincent B. Young
    affiliations:
      - ref: micro
      - ref: intmed
  - name: Krishna Rao
    affiliations:
      - ref: intmed
  - name: Patrick D. Schloss
    orcid: 0000-0002-6935-4275
    email: pschloss@umich.edu
    corresponding: true
    affiliations:
      - ref: micro
      - ref: ccmb
affiliations:
  - id: bioinf
    name: Department of Computational Medicine & Bioinformatics, University of Michigan
  - id: micro
    name: Department of Microbiology & Immunology, University of Michigan
  - id: mcdb
    name: Department of Molecular, Cellular, and Developmental Biology, University of Michigan
  - id: intmed
    name: Division of Infectious Diseases, Department of Internal Medicine, University of Michigan
  - id: ccmb
    name: Center for Computational Medicine and Bioinformatics, University of Michigan
abstract: TODO
importance: TODO
keywords: _C. difficile_ infection, supervised machine learning
---

@@

@@

TODO [@schloss_introducing_2009;@topcuoglu_mikropml_2021;@sovacool_schtools_2022]

@@

>>>>

---
title: "Predicting _C. difficile_ Infection Severity from the Taxonomic Composition of the Gut Microbiome"
bibliography: references.bib
fontsize: 11pt
date: last-modified
date-format: medium
format:
  manuscript-pdf:
    mainfont: Helvetica
    keep-tex: true
    linenumbers: true
    doublespacing: true
    lof: false
  html:
    embed-resources: true
runningtitle: Predicting Severe _C. difficile_ infections
runningauthor: Sovacool
papertype: Preprint
author:
  - name: Kelly L. Sovacool
    orcid: 0000-0003-3283-829X
    affiliations:
      - ref: bioinf
  - name: Sarah E. Tomkovich
    affiliations:
      - ref: micro
  - name: Megan L. Coden
    affiliations:
      - ref: mcdb
  - name: Vincent B. Young
    affiliations:
      - ref: micro
      - ref: intmed
  - name: Krishna Rao
    affiliations:
      - ref: intmed
  - name: Patrick D. Schloss
    orcid: 0000-0002-6935-4275
    email: pschloss@umich.edu
    corresponding: true
    affiliations:
      - ref: micro
      - ref: ccmb
affiliations:
  - id: bioinf
    name: Department of Computational Medicine & Bioinformatics, University of Michigan
  - id: micro
    name: Department of Microbiology & Immunology, University of Michigan
  - id: mcdb
    name: Department of Molecular, Cellular, and Developmental Biology, University of Michigan
  - id: intmed
    name: Division of Infectious Diseases, Department of Internal Medicine, University of Michigan
  - id: ccmb
    name: Center for Computational Medicine and Bioinformatics, University of Michigan
abstract: TODO
importance: TODO
keywords: _C. difficile_ infection, supervised machine learning
---

@@

## Model performance

Report median AUROC for trainset and testset, and median AUBPRC for testset (@fig-performance).

## Feature importance

(@fig-features)

## Clinical value of severity prediction models

@@

## Sample collection and 16S rRNA gene amplicon sequencing

[@schloss_introducing_2009]

## Defining CDI severity

IDSA definition of severe CDI based on lab values.
CDC definiton of severe CDI based on disease-related complications [@mcdonald_recommendations_2007].

## Model training and evaluation

[@topcuoglu_mikropml_2021]

### Balanced precision

## Code and data availability

[@sovacool_schtools_2022]

@@

# References {.unnumbered}

::: {#refs}
:::

# Figures

::: {#fig-flowchart fig-cap="**CDI severity definitions.**
A) Decision flow chart to define CDI cases as severe according to the Infectious Diseases Society of America (IDSA) based on lab values,
the occurence of complications due to any cause (All-cause),
and the occurence of disease-related complications confirmed as attributable to CDI with chart review (Attrib).
B) The proportion of severe CDI cases labelled according to each definition.
An additional 'Pragmatic' severity definition uses the Attributable definition when possible,
and falls back to the All-cause definition when chart review is not available.
"}
![](figures/flowchart_sankey.png)
:::

::: {#fig-performance fig-cap="**Performance of ML models.**
Area Under the Receiver-Operator Characteristic Curve (AUROC) for the cross-validation trainsets and testsets,
and the Area Under the Balanced Precision-Recall Curve (AUPRC) for the testsets.
Left: models were trained on the full dataset, with different numbers of samples available for each severity definition.
Right: models were trained on the intersection of samples with all labels available for each definition.
"}
![](figures/ml-performance.png)
:::

::: {#fig-features fig-cap="**Feature importance.**"}
TODO insert figure here
:::
